---
title: "Formatting Tables"
author: "Alan Arnholt"
output: 
    bookdown::html_document2:
      theme: yeti
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center")
```

## Tables and Output {-}

This document is a work in progress.  

```{r label = "cars"}
mod <- lm(dist ~ speed, data = cars)
# Just the coefficients
coef(summary(mod))
# Or
summary(mod)$coef
```

There are several packages one might use to format the output from a linear model.

* `knitr` and the function `kable()`
* `xtable` and the function `xtable()`
* `htmlTable` and the function `htmlTable()`

```{r, label = "test1"}
knitr::kable(coef(summary(mod)), caption = "Write your descriptive caption between the quotes", booktabs = TRUE, digits = 4)
```

```{r}
knitr::kable(coef(summary(mod)), caption = "Write your descriptive caption between the quotes 2", booktabs = TRUE, digits = 4, col.names = c("$\\alpha$", "$\\beta$", "$\\delta$", "$\\eta$"))
```


However, I can not figure out how to refer to the created Tables in a bookdown document with any package other than `knitr`.  If you can help, please do.



```{r, label = "motab"}
knitr::kable(mtcars[1:5, ],  booktabs = TRUE, caption = "Something here")
```

Table \@ref(tab:motab)!

```{r, label = "reg", results = 'asis'}
XT <- xtable::xtable(coef(summary(mod)), caption = "Write your descriptive caption between the quotes", auto = TRUE)
print(XT, type = "html")
```

```{r, label = "summTAB"}
options(table_counter = 3)
htmlTable::htmlTable(round(summary(mod)$coef, 4), css.cell = "padding-left: 1em; padding-right: 1em;", caption ="`dist` regressed on `speed`", align = "rrrr")
```

```{r, label = "DT", fig.cap = "Something new here please"}
DT::datatable(iris)
```

```{r, warning = FALSE}
pander::pander(summary(mod)$coef, caption = "You add it here")
```

```{r}
knitr::kable(head(iris), format = "html", table.attr = "id=\"mytable\"")
```

Table: (\#tab:foo) Your table caption.

| Sepal.Length| Sepal.Width| Petal.Length|
|------------:|-----------:|------------:|
|          5.1|         3.5|          1.4|
|          4.9|         3.0|          1.4|
|          4.7|         3.2|          1.3|
|          4.6|         3.1|          1.5|

Table \@ref(tab:foo)

```{r, label = "TRIAL"}
# options(table_counter = 3)
htmlTable::htmlTable(round(summary(mod)$coef, 4), css.cell = "padding-left: 1em; padding-right: 1em;", caption ="`dist` regressed on `speed`", align = "rrrr")
```

Look  at Table \@ref(tab:test1)

## For Indy {-}

```{r, label = "indy"}
mx <- data.frame(j = 1:8, n = seq(50, 400, length = 8), ENA1 = rexp(8), ESA1 = rexp(8), ENA2 = rexp(8), ESA2 = rexp(8))
mx <- as.matrix(mx)
mx <- round(mx, 7)
options(table_counter = 5)
htmlTable::htmlTable(mx,
      align = "lrllll",
      css.cell = "padding-left: 1em; padding-right: 1em;",
      ctable=c("single", "single"),
      header = c("j", "n", "ESA", "ENA", "ESA", "ENA"),
      cgroup =c("", "", "(n = 1000, k = 400, &alpha; = 0.3)", "(n = 1000, k = 300, &beta; = 0.8)"),
      n.cgroup = c(1, 1, 2, 2),
      col.rgroup = c("none", "#F7F7F7"),
      caption = "Table for Indy")
```

