---
title: "Test It First!"
author: |
  | [\textcolor{blue}{Alan T. Arnholt}](https://alanarnholt.github.io)
institute: |
  | [\textcolor{blue}{Department of Mathematical Sciences}](http://mathsci.appstate.edu)
  | [\textcolor{blue}{Appalachian State University}](http://appstate.edu)
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
output:
  beamer_presentation:
    theme: "Malmoe"
    colortheme: "seahorse"
    fonttheme: "professionalfonts"
---

```{r, label = "setup", include = FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.width = 4, fig.height = 3, comment = NA)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see [\textcolor{blue}{http://rmarkdown.rstudio.com}](http://rmarkdown.rstudio.com).

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets


- Bullet 1
\pause
- Bullet 2
\pause
- Bullet 3
\pause
- Bullet 4
\pause
* Something else


[\textcolor{blue}{My web page}](https://alanarnholt.github.io)

## Slide with R Output

```{r cars}
summary(cars)
```

## Slide with Plot (dev = "tikz")

```{r, label = "aplot", echo = FALSE, dev = "tikz"}
library(ggplot2)
ggplot(pressure, aes(x = temperature, y = pressure)) + 
  geom_point(color = "blue") + 
  theme_bw()
```

## Mathematics 

\begin{equation}
\bar{x} = \sum_{i=1}^{n}\frac{x_i}{n}
\end{equation}

\begin{equation}
s^2 = \sum_{i=1}^{n}\frac{(x_i - \bar{x})^2}{n-1}
\end{equation}