---
title: "Behind the Curtain"
author: "Alan T. Arnholt"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: 
  bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5, 
                      prompt = TRUE, highlight = TRUE, tidy = FALSE, warning = FALSE, 
                      message = FALSE)
```

# Showing Code Chunks^[This is a footnote for Section \@ref(SCC)]{#SCC} 


````markdown
`r ''````{r, label = "SETUP", echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5, 
                      prompt = TRUE, highlight = TRUE, tidy = FALSE, warning = FALSE, 
                      message = FALSE)
```
````

## Subsection (Second level header)^[See the [RStudio Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)]{#SLH}

```{r, label = "HIS", fig.cap = "Histogram of randomly generated data"}
set.seed(123)
junk <- rnorm(10000, 100, 10)
hist(junk, col = "lightblue", breaks = "scott", main = "", xlab = "")
```

Look at Figure \@ref(fig:HIS) which has a mean of `r mean(junk)` and a standard deviation of `r sd(junk)`.^[The mean and standard deviation of `junk` are computed using inline `R` code.]

To include literal backticks, use more backticks outside, e.g., you can use two backticks to preserve one backtick inside: ``` `` `code` `` ```.^[The inline `R` code to compute the mean and standard deviation of `junk` are written `` ``r ''`r mean(junk)` ``, and  `` ``r ''`r sd(junk)` ``.]

* * *

When in doubt, read [bookdown: Authoring Books and Technical Documents with R Markdown](https://bookdown.org/yihui/bookdown/)!

Click [here](https://raw.githubusercontent.com/alanarnholt/GeneralStatistics/master/rmarkdown/BehindTheCurtain.Rmd) for the source code for this document.