---
title: "Miscelaneous Tips"
date: 'Last compiled: `r format(Sys.time(), "%A, %B %d, %Y - %X.")`'
author: 'Alan T. Arnholt'
output: 
    bookdown::html_document2
---

```{r label = "setup", include = FALSE, message = FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center")
```

# Using $\LaTeX$ 

* Start and end inline mathematics with one `$`
* Start and end displayed mathematics with two `$$`
* For more complex environments, [click here.](https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics)
* [Symbol table](http://mirror.jmu.edu/pub/CTAN/info/examples/Math_into_LaTeX-4/SymbolTables.pdf)
* To bold mathematics, use `\mathbf{}`.  For example, to bold $\Sigma$ use `\mathbf{\Sigma}` $\rightarrow \mathbf{\Sigma}$
* To create a superscript with text, use `^{\text{Your text}}`.  For example, to write the $i^{\text{th}}$ example use `$i^{\text{th}}$`. 


# Labeling Equations with `bookdown`

> Equation labels must start with the prefix `eq:` in `bookdown`. All labels in `bookdown` must only contain alphanumeric characters, `:`, `-`, and/or `/`. Equation references work best for LaTeX/PDF output, and they are not well supported in Word output or e-books. For HTML output, `bookdown` can only number the equations with labels. Please make sure equations without labels are not numbered by either using the `equation*` environment or adding `\nonumber` or `\notag` to your equations. The same rules apply to other math environments, such as `eqnarray`, `gather`, `align`, and so on (e.g., you can use the `align*` environment).

To create Equation \@ref(eq:label1), use the code shown in the gray box below.


\begin{equation} 
 f(x)=(x+a)(x+b)
 (\#eq:label1)
\end{equation}


```{r, label = "ex1", eval = FALSE}
\begin{equation} 
 f(x)=(x+a)(x+b)
 (\#eq:label1)
\end{equation}
```

To refer to Equation \@ref(eq:label1), use `\@ref(eq:label1)`.

To create the aligned Equation \@ref(eq:aligned), use the code shown in the gray box below.

\begin{align}
 f(x) &= x^4 + 7x^3 + 2x^2 \nonumber \\
      &\qquad {} + 10x + 12
(\#eq:aligned)      
\end{align}

```{r, label = "ex2", eval = FALSE}
\begin{align}
 f(x) &= x^4 + 7x^3 + 2x^2 \nonumber \\
      &\qquad {} + 10x + 12
(\#eq:aligned)      
\end{align}
```

To write the piecewise function in Equation \@ref(eq:piece), use the code shown in the gray box below.

\begin{equation}
 u(x) = 
  \begin{cases} 
   \exp{x} & \text{if } x \geq 0 \\
   1       & \text{if } x < 0
  \end{cases}
  (\#eq:piece)
\end{equation}

```{r, label = "ex3", eval = FALSE}
\begin{equation}
 u(x) = 
  \begin{cases} 
   \exp{x} & \text{if } x \geq 0 \\
   1       & \text{if } x < 0
  \end{cases}
  (\#eq:piece)
\end{equation}
```

# Labeling Graphs with `bookdown` 

Figure \@ref(fig:HISTO) was created using the `R` code chunk and code below.  Note that to refer to Figure \@ref(fig:HISTO), the name of the code chunk label is used.  In this case, the code chunk label is `HISTO` and to refer to Figure \@ref(fig:HISTO) one uses the syntax `\@ref(fig:HISTO)`.


    ```{r, label = "HISTO", fig.cap = "Write your descriptive caption here", echo = FALSE}`r ''`
    set.seed(13)
    stuff <- rnorm(1000, 100, 15)
    hist(stuff, col = "pink", breaks = "Scott")
    ```



```{r, label = "HISTO", fig.cap = "Write your descriptive caption here", echo = FALSE}
set.seed(13)
stuff <- rnorm(1000, 100, 15)
hist(stuff, col = "pink", breaks = "Scott")
```

# Labeling Tables with `bookdown` 

Table \@ref(tab:FT) was created using the `R` code chunk and code below.  Note that to refer to Table \@ref(tab:FT), the name of the code chunk label is used.  In this case, the code chunk label is `FT` and to refer to Table \@ref(tab:FT) one uses the syntax `\@ref(tab:FT)`.

    ```{r, label = "FT", echo = FALSE}`r ''`
    knitr::kable(head(iris), booktabs = TRUE,  
    caption = 'The first six rows of `iris`'
    )
    ```
    
```{r, label = "FT", echo = FALSE}
knitr::kable(head(iris), booktabs = TRUE,  caption = 'The first six rows of `iris`')
```  



