---
title: "Relational Operators"
author: "Based on DataCamp"
date: '`r format(Sys.time(), "%b %d, %Y")`'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, prompt = TRUE)
```

## Relational Operators (1)

```{r}
TRUE == TRUE
TRUE == FALSE
TRUE != TRUE
TRUE != FALSE
```

## Relational Operators (2)

```{r}
"hello" == "hola"
5 == 2
"hello" != "hola"
5 != 2
```

## Relational Operators (3)

```{r}
2 < 5
2 > 5
"adios" > "goodbye"  # alphabetical order!
FALSE < TRUE         # FALSE coerces to 0 and TRUE coerces to 1
```

## Relational Operators (4)

```{r}
5 >= 2
5 >= 5
5 <= 5
```

## Relational Operators & Vectors 

```{r}
boone_temp <- c(81, 72, 73, 75, 69)
boone_temp
boone_temp > 72
boone_temp >= 72
```

## Logical Operators

* AND operator `&`
* OR operator `|`
* NOT operator `!`

## AND operator (`&`)

```{r}
TRUE & TRUE       # Only TRUE if both are TRUE
FALSE & TRUE      # FALSE if both not TRUE
TRUE & FALSE
FALSE & FALSE
```

## AND operator (`&`)

```{r}
x <- 15
x > 5 & x < 20    # BOTH conditions are TRUE
x > 5 & x < 14    # ONLY one condition is TRUE 
```

## OR operator (1)

```{r}
TRUE | TRUE   # TRUE if at least one is TRUE
TRUE | FALSE  # ONLY FALSE if both are FALSE
FALSE | TRUE
FALSE | FALSE
```

## OR operator (2)

```{r}
y <- 11
y < 15 | y > 15  # At least one TRUE -> TRUE
y < 10 | y > 12  # FALSE if both FALSE
```

## NOT operator (!)

```{r}
!TRUE
!FALSE
x
!(x <= 3)
```

## NOT operator (!)

```{r}
is.numeric(10)
!is.numeric(10)
is.numeric("hola")
!is.numeric("hola")
```

## Logical Operators & Vectors

```{r}
c(TRUE, TRUE, FALSE) & c(TRUE, FALSE, FALSE)
c(TRUE, TRUE, FALSE) | c(TRUE, FALSE, FALSE)
!c(TRUE, TRUE, FALSE)
```

## TEASER - Explain what is happening

```{r}
c(TRUE, TRUE, FALSE, TRUE, FALSE) & c(TRUE, FALSE, FALSE)
```

## Vector and (Scalar) Operators

```{r}
c(TRUE, TRUE, FALSE) & c(FALSE, FALSE, FALSE)
c(TRUE, TRUE, FALSE) && c(FALSE, FALSE, FALSE)
c(TRUE, TRUE, FALSE) | c(FALSE, FALSE, FALSE)
c(TRUE, TRUE, FALSE) || c(FALSE, FALSE, FALSE)
```

## `if` statement (1)

```{r, eval = FALSE}
if(condition){
  expression
}
```

```{r}
x <- 5
if(x >= 0){
  print("x is a positive number")
}
```

## `if` statement (2)

```{r, eval = FALSE}
if(condition){
  expression
}
```

```{r}
x <- 5
if(x < 0){                        # condition is FALSE
  print("x is a negative number") # Print not executed 
}                                 # No output!
```

## `else` statement (1)

```{r, eval = FALSE}
if(condition){
  expression1
} else {
  expression2
}
```

## `else` statement (2)

```{r, eval = FALSE}
if(condition){
  expression1
} else {
  expression2
}
```

```{r}
x <- -5
if(x < 0){   #Condition is TRUE
  print("x is a negative number")
} else {
  print("x is either a positive number or zero")  
}
```

## `else` statement (3)

```{r, eval = FALSE}
if(condition){
  expression1
} else {
  expression2
}
```

```{r}
x <- 5
if(x < 0){   #Condition is TRUE
  print("x is a negative number")
} else {
  print("x is either a positive number or zero")  
}
```

## `else if` statement {.smaller}

<div class="columns-2">
```{r}
x <- -3
if(x < 0) {
  print("x is a negative number")
} else if(x == 0) {
  print("x is zero")
} else {
  print("x is a positive number")
}
```

```{r, eval = FALSE}
# else if TEMPLATE
if(condition1) {
  expression1
} else if(condition2) {
  expression2
} else {
  expression3
}
```
</div>

## `else if` statement (2) {.smaller}

<div class="columns-2">
```{r}
x <- 0
if(x < 0) {
  print("x is a negative number")
} else if(x == 0) {
  print("x is zero")
} else {
  print("x is a positive number")
}
```

```{r, eval = FALSE}
# else if TEMPLATE
if(condition1) {
  expression1
} else if(condition2) {
  expression2
} else {
  expression3
}
```
</div>

## `else` statement {.smaller}

<div class="columns-2">
```{r}
x <- 5
if(x < 0) {         #Cond is FALSE
  print("x is a negative number")
} else if(x == 0) { #Cond is FALSE
  print("x is zero")
} else {
  print("x is a positive number")
}
```

```{r, eval = FALSE}
# TEMPLATE
if(condition1) {
  expression1
} else if(condition2) {
  expression2
} else {
  expression3
}
```
</div>



## `if`, `else if`, `else` {.smaller}

<div class="columns-2">
```{r}
x <- 9
if(x %% 2 == 0) {        
  print("divisible by 2")
} else if(x %% 3 == 0) { 
  print("divisible by 3")
} else {
  print("not divisible by 2 nor by 3...")
}
```

```{r, eval = FALSE}
# TEMPLATE
if(condition1) {
  expression1
} else if(condition2) {
  expression2
} else {
  expression3
}
```
</div>
