---
title: "Creating and Publishing a Web Page on GitHub"
date: 'Last compiled: `r format(Sys.time(), "%A, %B %d, %Y - %X.")`'
author: 'Alan T. Arnholt'
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, comment = NA)
```

## Prerequisites

This talk assumes you have the following installed on your machine:

* `R` (https://cran.r-project.org/) 
* `RStudio` (https://www.rstudio.com/)
* `Git` (https://git-scm.com/)

or that you have access to the ASU [RStudio server](https://mathr.math.appstate.edu/auth-sign-in) that has all of the above and more already installed.

**Note:** If you are a MAC user, I recommend using the version of `Git` that is pre-installed with Xcode (https://developer.apple.com/xcode/).  See the Mac OS section of [Happy Git and GitHub for the userR](http://happygitwithr.com/install-git.html#mac-os) for details.


## Tips/Checks

To see the version of `Git` installed on your machine, open a shell and type: 

```{r, label = "version", engine = "bash", echo = TRUE}
git --version
```

To find out where `Git` is installed on your machine, open a shell and type:

```{r, label = "which", engine = "bash", echo = TRUE}
which git
```

## Make Sure `Git` Knows Who You Are

Type the following at the shell prompt:

```{r label = "who", engine = "bash", echo = TRUE}
git config --global --list
```

If you have not configured at a minimum your `user.name` and `user.email`, substitute your name and email for mine in the code below.

```{r, label = "youme", engine = "bash", echo = TRUE, eval = FALSE}
git config --global user.name 'Alan Arnholt'
git config --global user.email 'arnholtat@appstate.edu'
```

      


## Create a GitHub Account

* Username advice

* Sign up (https://github.com/)

* Request a student discount (https://education.github.com/)
