---
title: "Creating and Publishing a Web Page on GitHub"
date: 'Last compiled: `r format(Sys.time(), "%A, %B %d, %Y - %X.")`'
author: 'Alan T. Arnholt'
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

## Prerequisites

This talk assumes you have the following installed on your machine:

* `R` (https://cran.r-project.org/) 
* `RStudio` (https://www.rstudio.com/)
* `Git` (https://git-scm.com/)

or that you have access to the ASU [RStudio server](https://mathr.math.appstate.edu/auth-sign-in) that has all of the above and more already installed.

**Note:** If you are a MAC user, we recommend using the version of `Git` that is pre-installed with Xcode (https://developer.apple.com/xcode/).


## MAC tips

To see the version of `Git` is installed on your machine, open a shell and type:

```{r, engine = "bash"}
git version
```

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

